Exploring dimensions of biodiversity in Japanese ferns

### 2021-08-24 ms/SI.Rmd
Update MS
fix maps in SI
move more ms functions to ms_functions.R

```{r envir, fig.width = 7, fig.height = 9}
# Define function for plotting environment
plot_env <- function(part, label, var_select, fill_lab = "", labels = ggplot2::waiver()) {
  ggplot() +
    geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
    geom_sf(data = biodiv_ferns_for_plotting, aes(fill = .data[[var_select]]), color = "transparent", alpha = 0.8) +
    annotate(
      geom="richtext", y = Inf, x = -Inf, 
      label = glue("<span style='color:black'>**({part})** {label}</span>"), 
      hjust = "inward", vjust = "inward", color = "white") + 
    labs(fill = fill_lab) +
    scale_fill_scico(palette = "lajolla", labels = labels) +
    scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
    theme_grey(base_size = 10) +
    theme(
      panel.grid.major = element_line(color = "grey60", size = 0.1),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "transparent"),
      axis.ticks = element_blank(),
      axis.text = element_text(color = "grey60"),
      axis.title = element_blank(),
      legend.position = "right",
      legend.title = element_markdown()
    )
}

@@

a <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = abundance), color = "transparent", alpha = 0.8) +
  scale_fill_scico(palette = "lajolla", label = k) +
  labs(fill = "No. specimens") +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)**</span>", hjust = "inward", vjust = "inward", color = "white") + 
  map_theme()

b <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = redundancy), color = "transparent", alpha = 0.8) +
  scale_fill_scico(palette = "lajolla") +
  labs(fill = "Redundancy") +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)**</span>", hjust = "inward", vjust = "inward", color = "white") + 
  map_theme() +
  theme(
    axis.text.y = element_blank()
  )

@@

```{r endemism-restricted, fig.width = 5, fig.height = 5}
ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_endemic_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
  scale_fill_manual(values = canape_cols) +
  map_theme() +
  theme(legend.title = element_blank())
```

@@

  ggplot() +
    geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
    geom_sf(data = genus_abun, aes(fill = richness), color = "transparent", alpha = 0.8) +
    annotate(
      geom="richtext", y = Inf, x = -Inf, 
      label = glue("<span style='color:black'>**({part})** *{genus_select}*</span>"), 
      hjust = "inward", vjust = "inward", color = "white") + 
    labs(fill = "Richness") +
    scale_fill_scico(palette = "lajolla") +
    scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
    theme_grey(base_size = 10) +
    theme(
      panel.grid.major = element_line(color = "grey60", size = 0.1),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "transparent"),
      axis.ticks = element_blank(),
      axis.text = element_text(color = "grey60"),
      axis.title = element_blank(),
      legend.position = "right",
      legend.title = element_markdown()
    )
}

>>>>

```{r envir, fig.width = 7, fig.height = 9}
# Define function for plotting environment
plot_env <- function(part, label, var_select, fill_lab = "", labels = ggplot2::waiver()) {
  ggplot() +
    geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
    geom_sf(data = biodiv_ferns_spatial, aes(fill = .data[[var_select]]), color = "transparent") +
    geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
    annotate(
      geom="richtext", y = Inf, x = -Inf, 
      label = glue("<span style='color:black'>**({part})** {label}</span>"), 
      hjust = "inward", vjust = "inward", color = "white") + 
    labs(fill = fill_lab) +
    scale_fill_scico(palette = "lajolla", labels = labels) +
    scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
    theme_grey(base_size = 10) +
    theme(
      panel.grid.major = element_line(color = "grey60", size = 0.1),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "transparent"),
      axis.ticks = element_blank(),
      axis.text = element_text(color = "grey60"),
      axis.title = element_blank(),
      legend.position = "right",
      legend.title = element_markdown()
    )
}

@@

a <- ggplot() + 
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = abundance), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  scale_fill_scico(palette = "lajolla", label = k) +
  labs(fill = "No. specimens") +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)**</span>", hjust = "inward", vjust = "inward", color = "white") + 
  map_theme()

b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = redundancy), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  scale_fill_scico(palette = "lajolla") +
  labs(fill = "Redundancy") +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)**</span>", hjust = "inward", vjust = "inward", color = "white") + 
  map_theme() +
  theme(
    axis.text.y = element_blank()
  )

@@

```{r endemism-restricted, fig.width = 5, fig.height = 5}
ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_endemic_spatial, aes(fill = endem_type), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  scale_fill_manual(values = canape_cols) +
  map_theme() +
  theme(legend.title = element_blank())
```

@@

  ggplot() +
    geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
    geom_sf(data = genus_abun, aes(fill = richness), color = "transparent") +
    geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
    annotate(
      geom="richtext", y = Inf, x = -Inf, 
      label = glue("<span style='color:black'>**({part})** *{genus_select}*</span>"), 
      hjust = "inward", vjust = "inward", color = "white") + 
    labs(fill = "Richness") +
    scale_fill_scico(palette = "lajolla") +
    scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
    theme_grey(base_size = 10) +
    theme(
      panel.grid.major = element_line(color = "grey60", size = 0.1),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "transparent"),
      axis.ticks = element_blank(),
      axis.text = element_text(color = "grey60"),
      axis.title = element_blank(),
      legend.position = "right",
      legend.title = element_markdown()
    )
}

### 2021-08-24 ms/manuscript.Rmd
Update MS
fix maps in SI
move more ms functions to ms_functions.R

All computational analyses were carried out in R v.`r getRversion()` [@RCoreTeam2020] unless otherwise stated. The `r pack("targets")` R package was used to maintain analysis workflow [@Landau2021].

@@

Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]. 
However, using this dataset alone for phylogeny suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils available for molecular dating. 
Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree. 
To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa). 
We conducted phylogenetic analysis with `r software("IQTREE")` [@Nguyen2015] and molecular dating with `r software("treePL")` [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020]. 
We then trimmed the tree to Japanese taxa only and used this for all further analyses.

>>>>

All computational analyses were carried out in R v`r getRversion()` [@RCoreTeam2020] unless otherwise stated. The `r pack("targets")` R package was used to maintain analysis workflow [@Landau2021].

@@

Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]. 
However, using this dataset alone for phylogeny suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils available for molecular dating. 
Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree. 
To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa). 
We conducted phylogenetic analysis with `r software("IQTREE")` v1.6.12 [@Nguyen2015] and molecular dating with `r software("treePL")` v1.0 [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020]. 
We then trimmed the tree to Japanese taxa only and used this for all further analyses.
